---
title: "7Shifts Data Analysis"
output:
  word_document: default
  html_document:
    df_print: paged
  pdf_document: default
---

## Objective
Find 1-3 interesting things you find about the conversion rates or other trends in the data.
Data provided consists of two files: companies and trials.

## Approach
- Load packages needed
- Load data (companies and trials)
- Get summary statistics for data
- Perform data cleaning if needed
- Find any trends or insights and add write-up as needed

## Load packages
```{r}
packages <- c('ggplot2', 'corrplot','tidyverse','scales','dplyr', 'lubridate','gridExtra','grid')
for (package in packages) {
  if (!require(package, character.only=T, quietly=T)) {
    install.packages(package)
    library(package, character.only=T)
  }
}
```

## Load data
```{r}
trials <- read_csv("trials.csv")
companies <- read_csv("companies.csv")
```

## Review data
```{r}
## Summary and preview company information
summary(companies)
glimpse(companies)
```

```{r}
## Summary and preview trial information
summary(trials)
glimpse(trials)
```

```{r}
## Drop first columns
trials[1] <- NULL
companies[1] <- NULL
```

## Business information
- Conversion rate is defined as the proportion of prospective customers (trials) that made
a purchase and became actual customers (companies) within a given time period.

- Trials can be joined to companies via company_id. A trial with no company ID did not
convert.

```{r}
# Combine trials and companies
trials_company <- trials %>%
  inner_join(companies,by = "company_id")
```
```{r}
## Create trial year
trials_company$trial_year <- lubridate::year(trials_company$trial_date)
## Unique years
unique(trials_company$trial_year)

```
```{r}
#filter for years on or after 2017 since there are some odd dates
trials_company <- trials_company %>%
  filter(trial_year >= 2017)
```

## Trends
```{r}
#trial year vs avg display price
trial_avg_display_price <- trials_company %>%
  dplyr::group_by(trial_year) %>%
  dplyr::summarise(avg_display_price = mean(display_price)) %>%
  dplyr::select(trial_year, avg_display_price)
  

ggplot(data=trial_avg_display_price, aes(x=(trial_year), y=avg_display_price)) +
  geom_bar(stat="identity", fill="steelblue")+ 
  theme_minimal()+
  xlab("Trial Year") + 
  ylab("Avg. Display Price")
```
- Looks like the average display price had a significant increase after 2017.  


```{r}
#trial year vs utm vs avg display price
trial_utm_avg_display_price <- trials_company %>%
  dplyr::group_by(trial_year, utm) %>%
  dplyr::summarise(avg_display_price = mean(display_price)) %>%
  dplyr::select(trial_year, utm, avg_display_price)

ggplot(trial_utm_avg_display_price, aes(x=(trial_year), y=avg_display_price, group = utm)) +
  geom_line(aes(color=utm)) +
  geom_point(aes(color=utm)) +
  theme_minimal()+
  xlab("Trial Year") + 
  ylab("Avg. Display Price")
```

- Organic average display price has been the highest price across all years.
- Google's average display price took a drop from 2018 and 2019 and has remained steady since.
- Linkedin's average display price also took a drop from 2018 to 2019 and has been steadily rising since.

```{r}
#differences between companies that have cancelled vs not cancelled
trial_utm_avg_display_price_other <- trials_company %>%
  dplyr::filter(is.na(cancellation_date)) %>%
  dplyr::group_by(trial_year, utm) %>%
  dplyr::summarise(avg_display_price = mean(display_price)) %>%
  dplyr::select(trial_year, utm, avg_display_price)

plot1 <- ggplot(trial_utm_avg_display_price_other, aes(x=(trial_year), y=avg_display_price, group = utm)) +
  geom_line(aes(color=utm)) +
  geom_point(aes(color=utm)) +
  theme_minimal()+
  ggtitle("Cancelled ")+
  xlab("Trial Year") + 
  ylab("Avg. Display Price")


trial_utm_avg_display_price_other1 <- trials_company %>%
  dplyr::filter(!is.na(cancellation_date)) %>%
  dplyr::group_by(trial_year, utm) %>%
  dplyr::summarise(avg_display_price = mean(display_price)) %>%
  dplyr::select(trial_year, utm, avg_display_price)

plot2 <- ggplot(trial_utm_avg_display_price_other1, aes(x=(trial_year), y=avg_display_price, group = utm)) +
  geom_line(aes(color=utm)) +
  geom_point(aes(color=utm)) +
  #geom_bar(stat="identity",position = "dodge")+ 
  theme_minimal()+
  ggtitle("Still active")+
  xlab("Trial Year") + 
  ylab("Avg. Display Price")


grid.arrange(plot1, plot2, ncol=2)

```

- For customers that are still active 
   - Linkedin data seems to have stopped after 2020.  Might be a data issue.  Needs more investigation
   - Google average display prices have been increasing with a dip between 2018 and 2020.  Had a sharp rise after 2020
   - Organic average display prices seem to be the most costly from 2017 to 2020.  Saw a drop after 2020


- For customers that have cancelled
   - Organic average display prices look to be the most expensive
   - Linkedin prices seem to be the lowest with mild fluctuations
   - Google average display prices have had mild fluctuations



